---
title: 'Exam Prep'
author: 'Phil 444'
date: 'April 12, 2024'
execute:
  warning: false
  message: false
  cache: true
format:
    html:
      toc: false
      mainfont: SF Pro Rounded
    pdf: 
      format: scrreprt
      keep-tex: true
      geometry:
        - top=20mm
        - left=25mm
        - heightrounded
      mainfont: Scala Pro
      sansfont: Scala Sans Pro
      mainfontoptions: Scale = MatchLowercase
      fontsize: 12pt
      linkcolor: black
      link-citations: true
---

- The actual exam will be 6 questions taken from the questions below.
- The numbers in the numerical questions will be changed, but the wording won't be.
- For any question asking for a verbal answer (e.g., explain this notion, or this constraint), you should answer in about **250** words. These are _short answer_ questions, not essays, and the aim is to either (a) simply describe a concept or idea, or (b) present one reason for or against some notion. Don't try to write a five paragraph essay for each one!

## Questions {-}

1. Briefly describe the difference between what Lackey calls "inflationary" and "deflationary" accounts of group belief, and give an example of each.
2. Explain the Conditionalisation constraint in Russell, Hawthorne and Buchak's paper _Groupthink_.
3. In a local election, with 100 voters, the voters have the following preferences. 45 people rank the candidates ADBC (i.e., A first, D second, B third, C fourth); 25 rank them BCDA; 20 rank them DABC; 10 rank them CDBA. Who would win if the voters vote sincerely, and the city uses first-past-the-post voting? Who would win if the voters vote sincerely, and the city uses ranked choice (i.e., alternate vote) voting? Very briefly (in a couple of sentences), which verdict do you think better reflects the will of the voters?
4. State Arrow's **Independence of Irrelevant Alternatives** condition.
5. Describe an example where Sen's Condition L (Liberalism) and Condition P (Pareto) conflict.
6. Consider an Axelrod-style Prisoners Dilemma tournament with just the following three strategies: Tit-for-Tat (i.e., C at move one, then do whatever the other person did last time); Grim Trigger (i.e., C until the other person plays D, then D forever); and a strategy that plays D until the other person plays D, then C forever. Each will play 100 rounds against the other two. The payoffs each round are (as normal), 5 points for playing D against C, 3 points for playing C against C, 1 point for playing D against D, 0 points for playing C against D. How many points over the 200 games (i.e., 100 rounds against 2 opponents) will each of them end up with?
7. Describe an example of a **focal point** in Schelling's sense.
8. What is the difference between weak dominance and strict dominance? Describe an example where the two notions come apart.
9. In @tbl-dom, what will be the result if both players use iterated deletion of strictly dominated strategies to decide what to do? (In each cell, Row's payouts are first, and Column's are second.)
10. In @fig-back, what will be the result if all players uses backward induction to solve the problem? (At each terminal node, the payouts are player 1's, then player 2's, then player 3's. Each player moves once; first player 1, then player 2, then player 3.)
11. In @fig-signal, describe a separating equilbrium of the game.
12. What is the human capital theory of the explanation of the college wage premium? Describe one objection to it. (You do not have to answer the objection.)

|      |   L  |   C   |   R   |
|:----:|:----:|:-----:|:-----:|
|   U  | 1,1  | 2,0   | 2,2   |
|   M  | 0,3  | 1,5   | 4,4   |
|   D  |  2,4 | 3,6   | 3,0   |

: Game table for Q9 {#tbl-dom}


```{r engine="tikz"}
#| fig-format: png
#| echo: false
#| label: fig-back
#| fig.cap: "Tree for Q10"


\usetikzlibrary{calc}

\begin{tikzpicture}
  \tikzset{
    % Three node styles for game trees: solid and hollow and square
      solid node/.style={circle,draw,inner sep=1.5,fill=black},
      hollow node/.style={circle,draw,inner sep=1.5},
      square node/.style={rectangle,draw, inner sep = 1, fill = black}
      }

  % Specify spacing for each level of the tree
  \tikzstyle{level 1}=[level distance=15mm,sibling distance=40mm]
  \tikzstyle{level 2}=[level distance=15mm,sibling distance=20mm]
  \tikzstyle{level 3}=[level distance=15mm,sibling distance=10mm]

  \node[hollow node,label=above:{P1}]{}
    child { node [solid node,label=left:{P2}] {}
      child { node [solid node,label=left:{P3}] {}
        child { node [] {3,1,1}
            edge from parent
            node[left]{L}}
        child { node [] {4,0,0}
            edge from parent
            node[right]{R}}
        edge from parent
            node[left]{L}
        }
      child { node [solid node,label=right:{P3}] {}
        child { node [] {1,3,0}
            edge from parent
            node[left]{L}}
        child { node [] {1,0,1}
            edge from parent
            node[right]{R}}
        edge from parent
            node[right]{R}
        }        
      edge from parent
        node[left]{L}
    }
    child { node [solid node,label=right:{P2}] {}
      child { node [solid node,label=left:{P3}] {}
        child { node [] {1,0,0}
            edge from parent
            node[left]{L}}
        child { node [] {2,1,1}
            edge from parent
            node[right]{R}}
        edge from parent
            node[left]{L}
        }
      child { node [solid node,label=right:{P3}] {}
        child { node [] {0,2,1}
            edge from parent
            node[left]{L}}
        child { node [] {1,0,2}
            edge from parent
            node[right]{R}}
        edge from parent
            node[right]{R}
        }        
      edge from parent
        node[right]{R}
    }
    
    ;
    \end{tikzpicture}
```

```{r engine='tikz'}
#| label: fig-signal
#| fig.ext: 'png'
#| cache: TRUE
#| echo: FALSE
#| fig.cap: "Tree for Q11"

\usetikzlibrary{calc}

\begin{tikzpicture}[scale=1.4,font=\footnotesize]
\tikzset{
% Two node styles for game trees: solid and hollow
solid node/.style={circle,draw,inner sep=1.5,fill=black},
hollow node/.style={circle,draw,inner sep=1.5},
square node/.style={rectangle,draw, inner sep = 1, fill = black}
}

% Specify spacing for each level of the tree
\tikzstyle{level 1}=[level distance=12mm,sibling distance=25mm]
\tikzstyle{level 2}=[level distance=15mm,sibling distance=15mm]
\tikzstyle{level 3}=[level distance=18mm,sibling distance=11mm]

% The Tree
\node(0)[hollow node,label=right:{Nature}]{}
child[grow=up]{node[solid node,label=above:{Proposer}] {}
child[grow=left]{node(1)[solid node]{}
child{node[square node,label=left:{1,2}]{} edge from parent node [above]{Up}}
child{node[square node,label=left:{1,0}]{} edge from parent node [below]{Down}}
edge from parent node [below]{L}
}
child[grow=right]{node(3)[solid node]{}
child{node[square node,label=right:{0,0}]{} edge from parent node [below]{Down}}
child{node[square node,label=right:{2,2}]{} edge from parent node [above]{Up}}
edge from parent node [below]{R}
}
edge from parent node [left, align=center]{A \\ 2/3}
}
child[grow=down]{node[solid node,label=below:{Proposer}] {}
child[grow=left]{node(2)[solid node]{}
child{node[square node,label=left:{2,0}]{} edge from parent node [above]{Up}}
child{node[square node,label=left:{0,2}]{} edge from parent node [below]{Down}}
edge from parent node [above]{L}
}
child[grow=right]{node(4)[solid node]{}
child{node[square node,label=right:{3,2}]{} edge from parent node [below]{Up}}
child{node[square node,label=right:{1,0}]{} edge from parent node [above]{Down}}
edge from parent node [above]{R}
}
edge from parent node [left,align=center]{B \\ 1/3}
};

% information set
\draw[dashed,rounded corners=10]($(1) + (-.55,.45)$)rectangle($(2) +(.55,-.45)$);
\draw[dashed,rounded corners=10]($(3) + (-.55,.45)$)rectangle($(4) +(.55,-.45)$);
% specify mover at 2nd information set
\node at ($(1)!.5!(2)$) {Respond};
\node at ($(3)!.5!(4)$) {Respond};
\end{tikzpicture}
```

